Finite Complete Suites for CSP Refinement Testing

### 2018-11-19 scp-journal/article/finitecomplete-failures.tex
Revised Section 3

\begin{lemma}\label{lemma:ufproperties}
If $s\in\trc(P)$ satisfies $\#s\le j$, then
$s, s.e\in\trc(U_F(j))$ for all $e\in\Sigma$, and the following properties hold.
\begin{eqnarray}
\label{eq:ifpa}
  &  & U_F(j)/s = U_F(j,\#s,G(P)/s)
\\
\label{eq:ifpb}
e\not\in [P/s]^0 & \implies & U_F(j)/s.e = (\efail\then\Stop)
\\
\label{eq:ifpc}
U_F(j)/s = U_F(j,\# s,n)  & \implies & [n]^0 = [P/s]^0
\\
\label{eq:ifpd}
U_F(j)/s = U_F(j,\# s,n)  & \implies & \minhits(n) = \minhits(P/s)
\end{eqnarray}
\end{lemma}
\begin{proof}
We prove (\ref{eq:ifpa}) by induction over the length of $s$. For $\#s = 0$,
the statement holds because $U_F(j)$ starts with the initial node $\ii n$ of
$G(P)$. Suppose that the statement holds for all traces $s$ with length $\# s
\le k < j$, so that $U_F(j)/s = U_F(j,\#s,G(P)/s)$. Now let $s.e$ be a trace
of $P$, so that $e\in [P/s]^0$. Since $[G(P)/s]^0 = [P/s]^0$ for all traces
$s$ of $P$, we conclude that $e\in  [G(P)/s]^0$, so $U_F(j,\#s,G(P)/s)$ can
engage into $e$ by executing branch (\ref{eq:ufc}). Since $t$ is the
transition function of $G(P)$ and $e\in [G(P)/s]^0$, $t(G(P)/s,e)$ is
defined, and $t(G(P)/s,e) = G(P)/s.e$. This leads to a new recursion, so that
$U_F(j)/s.e = U_F(j,\#s,G(P)/s)/e = U_F(j,\#s+1,G(P)/s.e)$ as required.

To prove (\ref{eq:ifpb}), we apply (\ref{eq:ifpa}) to conclude that $U_F(j)/s
= U_F(j,\#s,G(P)/s)$, because $s$ is a trace of $P$. Noting again that
$[G(P)/s]^0 = [P/s]^0$, this implies that $e\not\in [G(P)/s]^0$, so
$U_F(j,\#s,G(P)/s)$ can engage in $e$ by entering branch (\ref{eq:ufa}). The
specification of this branch implies that $_F(j)/s.e = U_F(j,\#s,G(P)/s)/e =
(\efail\then\Stop)$.

@@

\paragraph{Step~1.} Suppose that $P\lessdet_F Q$, so $P\lessdet_T Q$ and $Q\
conf\ P$ according to (\ref{eq:failconf}). Since   $\trc(Q)\subseteq
\trc(P)$, any adaptive test $U_F(j)$ running in parallel with $Q$ will always
enter the branches (\ref{eq:ufb}), (\ref{eq:ufc}), or (\ref{eq:ufd}) of the
external choice construction for $U_F(j,k,n)$. To see this, consider
 $U_F(j,k,n) = U_F(j)/s$
with $s\in\trc(Q)$. Lemma~\ref{lemma:ufproperties} implies $U_F(j,k,n) =
U_F(j,k,G(P)/s)$, so $[n]^0 = [G(P)/s]^0 = [P/s]^0$. As a consequence,
$[Q/s]^0\subseteq [P/s]^0 = [n]^0$, so branch~(\ref{eq:ufa}) can never be
entered in the parallel execution of $Q$ and $U_F(j)$, and the $fail$ event
cannot occur.

\paragraph{Step~2.} For proving that a test execution can never deadlock without a
$\epass$ event, it has to be shown that a test execution can neither block at
branch (\ref{eq:ufc}) nor at branch (\ref{eq:ufd}). These cases are
considered separately below.

>>>>

\begin{lemma}\label{lemma:ufproperties}
If $s\in\trc(P)$ satisfies $\#s\le j$, then
$s, s.e\in\trc(U_F(j))$ for all $e\in\Sigma$, and the following properties hold.
\begin{eqnarray}
\label{eq:ifpa}
  &  & U_F(j)/s = U_F(j,\#s,G(P)/s)
\\
\label{eq:ifpb}
e\not\in [P/s]^0 & \implies & U_F(j)/s.e = (\efail\then\Stop)
\\
\label{eq:ifpc}
U_F(j)/s = U_F(j,\# s,n)  & \implies & [n]^0 = [P/s]^0
\\
\label{eq:ifpd}
U_F(j)/s = U_F(j,\# s,n)  & \implies & \minhits(n) = \minhits(P/s)
\end{eqnarray}
\end{lemma}
\begin{proof}
We prove (\ref{eq:ifpa}) by induction over the length of $s$. For $\#s = 0$,
the statement holds because $U_F(j)$ starts with the initial node $\ii n$ of
$G(P)$. Suppose that the statement holds for all traces $s$ with length $\# s
\le k < j$, so that $U_F(j)/s = U_F(j,\#s,G(P)/s)$. Now let $s.e$ be a trace
of $P$, so that $e\in [P/s]^0$. Since $[G(P)/s]^0 = [P/s]^0$ for all traces
$s$ of $P$, we conclude that $e\in  [G(P)/s]^0$, so $U_F(j,\#s,G(P)/s)$ can
engage into $e$ by executing branch (\ref{eq:ufc}). Since $t$ is the
transition function of $G(P)$ and $e\in [G(P)/s]^0$, $t(G(P)/s,e)$ is
defined, and $t(G(P)/s,e) = G(P)/s.e$. So, the new recursion in branch
(\ref{eq:ufc}) is so that $U_F(j)/s.e = U_F(j,\#s,G(P)/s)/e =
U_F(j,\#s+1,G(P)/s.e)$ as required.

To prove (\ref{eq:ifpb}), we apply (\ref{eq:ifpa}) to conclude that $U_F(j)/s
= U_F(j,\#s,G(P)/s)$, because $s$ is a trace of $P$. Noting again that
$[G(P)/s]^0 = [P/s]^0$, this implies that $e\not\in [G(P)/s]^0$, so
$U_F(j,\#s,G(P)/s)$ can engage in $e$ by entering branch (\ref{eq:ufa}). The
specification of this branch implies that $U_F(j)/s.e = U_F(j,\#s,G(P)/s)/e =
(\efail\then\Stop)$.

@@

\paragraph{Step~1} Suppose that $P\lessdet_F Q$, so $P\lessdet_T Q$ and $Q\
conf\ P$ according to (\ref{eq:failconf}). Since   $\trc(Q)\subseteq
\trc(P)$, any adaptive test $U_F(j)$ running in parallel with $Q$ will always
enter the branches (\ref{eq:ufb}), (\ref{eq:ufc}), or (\ref{eq:ufd}) of the
external choice construction for $U_F(j,k,n)$. To see this, consider
 $U_F(j,k,n) = U_F(j)/s$
with $s\in\trc(Q)$. Lemma~\ref{lemma:ufproperties} implies $U_F(j,k,n) =
U_F(j,k,G(P)/s)$, so $[n]^0 = [G(P)/s]^0 = [P/s]^0$. As a consequence,
$[Q/s]^0\subseteq [P/s]^0 = [n]^0$, so branch~(\ref{eq:ufa}) can never be
entered in the parallel execution of $Q$ and $U_F(j)$, and the $fail$ event
cannot occur.

\paragraph{Step~2} For proving that a test execution can never deadlock without a
$\epass$ event, it has to be shown that a test execution can neither block at
branch (\ref{eq:ufc}) nor at branch (\ref{eq:ufd}). These cases are
considered separately below.

### 2018-11-19 scp-journal/article/preliminaries.tex
Revised Section 3

  & Property of $r_P, r_Q$ and (\ref{eq:refMaxRef})

@@

% -----------------------------------------------------------------------------
\paragraph{Minimal Hitting Sets of Minimal Acceptances} In this article, we are
interested in the minimal hitting sets of minimal acceptances; for these, the
abbreviated notation
$%\[ \
minhits(P/s) = \minhits(\minaccs(P/s))
$ %\]
is used. The minimal hitting sets
of minimal acceptances may be alternatively characterised by means of the
failures of a process as is done in~\cite{DBLP:conf/icfem/CavalcantiG07}. To
this end, in~\cite{DBLP:conf/icfem/CavalcantiG07}, the authors define, for
any collection ${\cal C}\subseteq 2^\Sigma$ of subsets from $\Sigma$
\begin{equation}\label{eq:minsub}
\minsub({\cal C}) = \{ A\in {\cal C}~|~\forall B\in {\cal C}:
B\subseteq A\Rightarrow B = A \}.
\end{equation}
The collection $\minsub({\cal C})$ contains all those sets of
${\cal C}$ that are not true subsets of other members of ${\cal C}$. With
this definition, the relation between failures and minimal hitting sets of
minimal acceptances is established in the following lemma.
%
\begin{lemma}
\label{lemma:failureshittingsets}
For any trace $s$ of CSP process $P$, define
$%\[
{\cal A}_s = \{ A\subseteq \Sigma~|~(s,A)\not\in\fails(P)  \}.
$ %\]
Then
$ %\[
\minhits(P/s) = \minsub {\cal A}_s
$ %\]
for all traces $s$ of $P$.
\end{lemma}
\begin{proof}
We derive
\begin{argue}
(s,A)\not\in {\text{failures}}(P)
\\
\Leftrightarrow A\not\in {\text{Ref}}(P/s)
\\
\Leftrightarrow  \forall R\in {\text{maxRef}}(P/s): A\not\subseteq R
\\
\Leftrightarrow   \forall B\in {\text{minAcc}}(P/s): A\not\subseteq (\Sigma\setminus B)& this follows from (\ref{eq:accref})
\\
\Leftrightarrow  \forall B\in {\text{minAcc}}(P/s): A\cap B\neq \varnothing
\\
\Leftrightarrow  A \text{ is a (not necessarily minimal)}
              \text{ hitting set of } {\text{minAcc}}(P/s)
\end{argue}
%
Specialising this derivation valid for arbitrary $A$ satisfying $(s,A)\not\in {\text{failures}}(P)$ to
minimal $A$ with this property proves the statement of the lemma.
\xbox
\end{proof}

@@

To prove ``$(2) \Rightarrow (1)$'', we assume that (2) holds, but that $P\
conf\ Q$ does {\it not} hold. According to Lemma~\ref{lemma:tgtrcref},
(\ref{eq:failrefb}), there exists $s\in\trc(P)\cap \trc(Q)$ such that
\[
\exists A_Q\in\minaccs(G(Q)/s): \forall A_P\in\minaccs(G(P)/s):
A_P\not\subseteq A_Q \qquad (*)
\]
Let $A$ be such an acceptance set $A_Q$ fulfilling (*).
Define
$%\[
\overline H = \bigcup\{ A_P \setminus A~|~A_P\in\minaccs(G(P)/s) \}.
$. %\]
Since $A_P \setminus A \neq\varnothing$ for all $A_P$ because of (*),
$\overline H$ is a hitting set of $\minaccs(G(P)/s)$ which has an  empty
intersection with $A$. Minimising $\overline H$ yields   a minimal hitting
set $H\in \minhits(P/s)$ which is {\it not} a hitting set of
$\minaccs(G(Q)/s)$, a contradiction to Assumption~2.
This completes the proof of the lemma.
\xbox
\end{proof}
%
We note that $\minaccs(P) = \{ \varnothing \}$ if $P = Q\intchoice \Stop$.
Since $\Stop$ accepts nothing, its minimal acceptance is $\varnothing$, and
this carries over to $Q\intchoice \Stop$.  From (\ref{eq:failrefb}) we
conclude that $\varnothing\in\minaccs(P)$ implies $\minaccs(P) = \{
\varnothing \}$. This clarifies that $\minhits(P/s)$ is empty if, and only
if, $\minaccs(P) = \{ \varnothing \}$. The proof of Lemma~\ref{lemma:hseta}
covers the situations where $\minaccs(P/s) = \{ \varnothing \}$ and so
$\minhits(P/s) = \varnothing$. Trivially,
\begin{equation}
\label{eq:minhitminaccempty}
\minaccs(P/s) = \{ \varnothing \} \iff \minhits(P/s) = \varnothing
\end{equation}
holds.

>>>>

  & Property of $r_P, r_Q$~(subset closure) and (\ref{eq:refMaxRef})

@@

% -----------------------------------------------------------------------------
\paragraph{Minimal Hitting Sets of Minimal Acceptances} In this article, we are
interested in the minimal hitting sets of minimal acceptances; for these, the
abbreviated notation
$%\[ \
\minhits(P/s) = \minhits(\minaccs(P/s))
$ %\]
is used. The minimal hitting sets
of minimal acceptances may be alternatively characterised by means of the
failures of a process as is done in~\cite{DBLP:conf/icfem/CavalcantiG07}. To
this end, in~\cite{DBLP:conf/icfem/CavalcantiG07}, the authors define, for
any collection ${\cal C}\subseteq 2^\Sigma$ of subsets from $\Sigma$
\begin{equation}\label{eq:minsub}
\minsub({\cal C}) = \{ A\in {\cal C}~|~\forall B\in {\cal C}:
B\subseteq A\Rightarrow B = A \}.
\end{equation}
The collection $\minsub({\cal C})$ contains all those sets of
${\cal C}$ that are not true subsets of other members of ${\cal C}$. With
this definition, the relation between failures and minimal hitting sets of
minimal acceptances is established in the following lemma.
%
\begin{lemma}
\label{lemma:failureshittingsets}
For any trace $s$ of CSP process $P$, define
$%\[
{\cal A}_s = \{ A\subseteq \Sigma~|~(s,A)\not\in\fails(P)  \}.
$ %\]
Then
$ %\[
\minhits(P/s) = \minsub {\cal A}_s
$ %\]
for all traces $s$ of $P$.
\end{lemma}
\begin{proof}
We derive
\begin{argue}
(s,A)\not\in \fails(P)
\\
\Leftrightarrow A\not\in Ref(P/s)
\\
\Leftrightarrow  \forall R\in maxRef(P/s): A\not\subseteq R
\\
\Leftrightarrow   \forall B\in minAcc(P/s): A\not\subseteq (\Sigma\setminus
B)& this follows from (\ref{eq:accref})
\\
\Leftrightarrow  \forall B\in minAcc(P/s): A\cap B\neq \varnothing
\\
\Leftrightarrow  A \text{ is a (not necessarily minimal)}
              \text{ hitting set of } minAcc(P/s)
\end{argue}
%
Specialising this derivation valid for arbitrary $A$ satisfying $(s,A)\not\in {\text{failures}}(P)$ to
minimal $A$ with this property proves the statement of the lemma.
\xbox
\end{proof}

@@

To prove ``$(2) \Rightarrow (1)$'', we assume that (2) holds, but that $P\
conf\ Q$ does {\it not} hold. According to Lemma~\ref{lemma:tgtrcref},
(\ref{eq:failrefb}), there exists $s\in\trc(P)\cap \trc(Q)$ such that
\[
\exists A_Q\in\minaccs(G(Q)/s): \forall A_P\in\minaccs(G(P)/s):
A_P\not\subseteq A_Q \qquad (*)
\]
Let $A$ be such an acceptance set $A_Q$ fulfilling (*).
Define
$%\[
\overline H = \bigcup\{ A_P \setminus A~|~A_P\in\minaccs(G(P)/s) \}.
$ %\]
Since $A_P \setminus A \neq\varnothing$ for all $A_P$ because of (*),
$\overline H$ is a hitting set of $\minaccs(G(P)/s)$ which has an  empty
intersection with $A$. Minimising $\overline H$ yields   a minimal hitting
set $H\in \minhits(P/s)$ which is {\it not} a hitting set of
$\minaccs(G(Q)/s)$, a contradiction to Assumption~2.
This completes the proof of the lemma.
\xbox
\end{proof}
%
We note that $\minaccs(P) = \{ \varnothing \}$ if $P = Q\intchoice \Stop$.
Since $\Stop$ accepts nothing, its minimal acceptance is $\varnothing$, and
this carries over to $Q\intchoice \Stop$.  From (\ref{eq:failrefb}) we
conclude that $\varnothing\in\minaccs(P)$ implies $\minaccs(P) = \{
\varnothing \}$. This clarifies that $\minhits(P/s)$ is empty if, and only
if, $\minaccs(P) = \{ \varnothing \}$. The proof of Lemma~\ref{lemma:hseta}
covers the situations where $\minaccs(P/s) = \{ \varnothing \}$ and so
$\minhits(P/s) = \varnothing$. Trivially,
\begin{equation}
\label{eq:minhitminaccempty}
\minaccs(P/s) = \{ \varnothing \} \iff \minhits(P/s) = \varnothing
\end{equation}
holds.
